State of the Practice for Lattice Boltzmann Method Software

### 2022-01-24 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits to Sections on identifying and filtering the software - answers RQ1

The above steps depend on interaction with a domain expert partner, as discussed
in Section~\ref{sec_vet_software_list}.

@@

Part of how we answer \rqref{RQ_WhatProjects}.  The candidate software was found
through search engine queries targeting authoritative lists of software. We
found LBM software listed on the websites GitHub and swMATH, as well as through
articles found in scholarly journals and databases.  The Domain Expert
(Section~\ref{sec_vet_software_list}) was also engaged in selecting the
candidate software.  \wss{Packages that fit our criteria but were found after
data collection and analysis was conducted, such as Musubi
\citep{HasertEtAl2014}, will need to be added to future SOP assessments.}

@@

\begin{enumerate}
	\item The software functionality must fall within the identified domain.
	\item The source code must be viewable.
	\item The empirical measures should be available, which implies a preference
	for GitHub-style repositories.
	\item The software cannot be marked as incomplete or in an initial
	development phase.
\end{enumerate}

The initial list had 45 packages, including a few packages that were later found
to not have publicly available source code, or to be in an incomplete state of
development.

@@

With previous section, how we answer \rqref{RQ_WhatProjects}.  Answer to
\rqref{RQ_WhatProjects} provided in this section.  To reduce the number of
members in the candidate software list to a manageable size, the following
filters were applied. The filters were applied in the priority order listed.

@@

While the initial list had 46 packages, filtering by scope, usage, and age
decreased the size of the list to 24 packages. Many of the 22 packages that were
removed could not be tested as there was no installation guide, they were
incomplete, source code was not publicly available, a license was needed, or the
project was out of scope or not up to a standard that would support
incorporating them into this study. These eliminated software packages are
listed in the Appendix of \citet{Michalski2021}. Of the remaining 24 packages
that were studied, some were kept on the list despite being marked as dead due
to their prevalence on authoritative lists on LBM software and due to their
surface excellence.  The final list of software packages that were analyzed in
this project can be found in the following two tables. Table~\ref{alivepackages}
lists packages that fell into the `alive' category as of mid 2020, and
Table~\ref{deadpackages} lists packages that were `dead' at that time.

@@

There is considerable variation among these software packages, including their
intended purpose, size, user interfaces, and software languages used. For
example, the OpenLB software package is predominantly a \CC{} package that makes
use of hybrid parallelization and was designed to address a range of CFD
problems \citep{heuveline2009towards}. The software package pyLBM is an
all-in-one Python language package for numerical simulations
\citep{graille2017pylbm}. ESPResSo is an extensible simulation package that is
specifically for research on soft matter, and is written in \CC{} and Python
\citep{weik2019espresso}. The HemeLB package is used for efficient simulation of
fluid flow in several medical domains, and is written predominantly in C, \CC,
and Python \citep{mazzeo2008hemelb}.

>>>>

The above steps depend on interaction with a Domain Expert partner, as discussed
in Section~\ref{sec_vet_software_list}.

@@

To answer~\rqref{RQ_WhatProjects} we needed to identify existing LBM projects.
The candidate software was found through search engine queries targeting
authoritative lists of software. We found LBM software listed on
\href{https://github.com/} {GitHub} and \href{https://swmath.org/} {swMATH}, as
well as through scholarly articles. The Domain Expert
(Section~\ref{sec_vet_software_list}) was also engaged in selecting the
candidate software.

@@

\begin{enumerate}
	\item The software functionality must fall within the identified domain.
	\item The source code must be viewable.
	\item The empirical measures should be available, which implies a preference
	for GitHub-style repositories.
	\item The software cannot be marked as incomplete or in an initial state of
	development.
\end{enumerate}

The initial list had 45 packages, including a few that were later found to not
have publicly available source code, or to be incomplete projects.

@@

To reduce the length of the list of software packages and to
answer~\rqref{RQ_WhatProjects}, we filter the initial list of 45.  The following
filters were applied in the priority order listed.

@@

Filtering by scope, usage, and age decreased the size of the list to 23
packages. Many of the 22 packages that were removed could not be tested as there
was no installation guide, they were incomplete, source code was not publicly
available, a license was needed, or the project was out of scope or not up to a
standard that would support incorporating them into this study. These eliminated
software packages are listed in the Appendix of \citet{Michalski2021}. Of the
remaining 23 packages that were studied, some were kept on the list despite
being marked as dead due to their prevalence on authoritative lists on LBM
software and due to their surface excellence.  After the initial measurement
exercise, which took place in mid 2020, we added one more package in January
2022: Musubi.  Musubi did not come up in our initial search, but once we learned
of it, we felt that this report is incomplete without including it.

The final list of 24 software packages that were analyzed in this project can be
found in two tables. Table~\ref{alivepackages} lists packages that fell into the
`alive' category as of mid 2020 (January 2022 for Musubi), and
Table~\ref{deadpackages} lists packages that were `dead' at that time. The
tables include a hyperlink to each project and, when available, citations for
relevant publications.  The final list of software packages shows considerable
variation in purpose, size, user interfaces, and programming languages. For
example, the OpenLB software package is predominantly a \CC{} package that makes
use of hybrid parallelization and was designed to address a range of CFD
problems \citep{heuveline2009towards}. The software package pyLBM is an
all-in-one Python language package for numerical simulations
\citep{graille2017pylbm}. ESPResSo is an extensible simulation package that is
specifically for research on soft matter, and is written in \CC{} and Python
\citep{weik2019espresso}. The HemeLB package is used for efficient simulation of
fluid flow in several medical domains, and is written predominantly in C, \CC,
and Python \citep{mazzeo2008hemelb}.

@@
